### 2020-02-24 chapters/background.tex
Background edit. Two-Stacks.

\section{Aggregations over data streams} \label{sec:aggregations}

An aggregation --- also known as a fold, reduce or accumulate --- is an operation that reduces a collection of values to a single one. Examples of such are the computation of data set sums, minima, averages or the number of distinct elements (cardinality). An aggregation may be exact or approximate. Exact aggregations compute a value from a sequence of elements, where no error or loss of resolution exists. On the other hand, approximate aggregations have an error margin to the associated aggregation value which is usually controllable by adjusting a set of parameters. Approximate aggregations trade precision of results for efficient time and space usage. Such trade-off between accuracy and time/space efficiency will be made clear when discussing some of the existing approximate aggregators in Section \ref{sec:pds}.

@@

@@

Different aggregations have different properties. These properties are listed in \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms}, where the authors group different aggregations according to the properties they satisfy. Table \ref{tbl:aggregations-properties} was adapted from this work and categorizes different families of operations based on said properties.

\begin{table}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/aggregation-operations-properties.png}
      \caption[Aggregation properties]{Aggregation properties. Checkmarks (\checkmark), crosses ($\times$), and question marks (?) indicate a property is true for all, false for all, or false for some of a given group of like operations, respectively}
      \label{tbl:aggregations-properties}
    \end{center}
\end{table}

@@

The following are formal definitions of these properties and were summarized from \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms}. An aggregation operation is said to have:

@@

%These properties are used to evaluate SWAG algorithms. Different algorithms might only apply to certain families of aggregations. This way, the multiple SWAG algorithms presented throughout this Thesis will be analyzed not only on their time and space complexity but also on the restrictions they impose regarding the aggregation' properties required.

Unlike many other scenarios that focus on data at rest, streaming aggregations handle data in motion. While some aggregations can be computed on an infinite data set, often we want to compute these over a finite one. As we have seen in Section \ref{sec:windows}, to obtain a bounded data set from a data stream we apply windowing techniques. In doing so, we obtain a finite data set denoted as a sliding window and compute aggregations over it. The authors of \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms} define such aggregations over sliding windows as Sliding Window Aggregations (SWAGs). Once again let us consider the data center temperature monitoring problem from Section \ref{sec:windows} as an example use-case for SWAGs. How would we compute the average temperature per rack in the data center in the last 5 minutes? To better understand, let us translate this question to a streaming SQL query, with a similar syntax to the API provided by Apache Flink \cite{flink}. Assuming all the temperatures can be queried from the \textit{Temperatures} table and that each rack corresponds a different \textit{rack\_id} we have the following query:

A sliding window aggregation (SWAG) algorithm allows the computation of an exact or approximate aggregation over a sliding window. A SWAG algorithm may restrict the set of possible to compute aggregations based on their properties. Different SWAG algorithms use different data structures for the window contents and the incremental aggregation state.

The aggregation properties analyzed in Section \ref{sec:agg-properties} are used to evaluate SWAG algorithms. Different algorithms might only apply to certain families of aggregations. This way, the multiple SWAG algorithms presented throughout this Thesis will be analyzed not only on their time and space complexity but also on the restrictions they impose regarding the aggregation' properties required.

Tangwongsan et al. present two of the most basic SWAG algorithms: \textit{Recalculate From Scratch (RFS)} and \textit{Subtract On Evict (SOE)}.

\subsubsection{Recalculate From Scratch}
Recalculate From Scratch (RFS) is the simplest and most general SWAG algorithm. As the name suggests, this algorithm recomputes the aggregation over the entire window for each insertion and eviction. For instance, the computation of a sum over a sliding window using RFS means that upon each insertion or eviction all elements are summed.

This approach does not restrain the aggregation in terms of properties: it works for every family of aggregations. However, given \textit{n} as the window size, this algorithm takes \textit{O(n)} space and time complexity which makes it unfit in most scenarios of large scale data processing.

\subsubsection{Subtract On Evict}
Subtract On Evict (SOE) only works for aggregations whose  \textit{combine} $\oplus$ function is invertible. That is, as defined in Table \ref{tbl:aggregations-properties}, a function $\ominus$ exists such that $(x \oplus y) \ominus y = x$.

SOE inserts new elements into the window and combines the previous aggregation value with the new element. On eviction, SOE applies the inverse of the combine function, $\ominus$, to expire the evicted element from the aggregation state.

With window size of \textit{n}, SOE has \textit{O(n)} space complexity and \textit{O(1)} time complexity. However, aggregations are not always invertible thus making this algorithm very restrictive and not always applicable.

>>>>

\section{Aggregations over data} \label{sec:aggregations}

An aggregation --- also known as a fold, reduce or accumulate --- is an operation that reduces a collection of values to a single one. Examples of such are the computation of data set sums, minima, averages or the number of distinct elements (cardinality). An aggregation may be exact or approximate. Exact aggregations compute a value from a sequence of elements, where no error or loss of resolution exists. On the other hand, approximate aggregations have an error margin to the associated aggregation value which is usually controllable by adjusting a set of parameters. 
Approximate aggregations are required when memory is limited, the volume of data to process is large and/or computing the exact aggregation can not be done in useful time. Thus, approximate aggregators trade precision of results for efficient time and space usage. In Section \ref{sec:pds} we explore approximate aggregators that make such a trade-off.

@@

While some aggregations can be computed on an infinite data set, often we want to compute these over a finite one. As we have seen in Section \ref{sec:windows}, to obtain a bounded data set from a data stream we apply windowing techniques. In doing so, we obtain a finite data set denoted as a sliding window and compute aggregations over it. The authors of \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms} define such aggregations over sliding windows as Sliding Window Aggregations (SWAGs). Considering the data center temperature monitoring problem from Section \ref{sec:windows} as an example use-case for SWAGs. How would we compute the average temperature per rack in the data center in the last 5 minutes? To better understand, let us translate this question to a streaming SQL query, with a similar syntax to the API provided by Apache Flink \cite{flink}. Assuming all the temperatures can be queried from the \textit{Temperatures} table and that each rack corresponds a different \textit{rack\_id} we have the following query:

A sliding window aggregation algorithm allows the computation of an exact or approximate aggregation over a sliding window. A SWAG algorithm may restrict the set of computable aggregations based on their properties. Different SWAG algorithms use different data structures for the window contents and the incremental aggregation state.

Two of the most basic SWAG algorithms are \textit{Recalculate-From-Scratch (RFS)} and \textit{Subtract-On-Evict (SOE)}.

Recalculate-From-Scratch (RFS) is the simplest and most general SWAG algorithm. As the name suggests, this algorithm recomputes the aggregation over the entire window for each window update. For instance, the computation of a sum over a sliding window using RFS means that whenever the window changes all elements are summed. This approach works for every aggregation. However, given \textit{n} as the window size, this algorithm takes \textit{O(n)} space and time complexity which makes it unfit in most scenarios of large scale data processing.

Subtract-On-Evict (SOE) on the other hand, is a SWAG algorithm that incrementally computes the aggregation value. For each element insertion or eviction from the window, the aggregation value is updated, in contrast to RFS which would recompute the entire aggregation. SOE relies on expiring the evicted elements from the aggregation state. Hence, SOE only works for \textit{invertible} aggregations. To illustrate, consider the computation of a sum aggregation over a sliding window. The total sum of the window will be the aggregation state or value. When inserting an element, the updated total window sum equals the total sum plus the new element. The sum aggregation is considered \textit{invertible} because, when evicting an element, the new aggregation state is the total sum computed this far minus the evicted element. With window size of \textit{n}, SOE has \textit{O(n)} space complexity and \textit{O(1)} time complexity. However, aggregations are not always invertible thus making this algorithm very restrictive and not always applicable.

Subtract-On-Evict is an improvement over the Recalculate-From-Scratch algorithm that  based on the idea of aggregation invertibility. The question then becomes if there are other aggregation properties besides invertibility that must be taken into account when analyzing general stream processing algorithms.

@@

Tangwongsan et al. construct Table \ref{tbl:aggregations-properties} to summarize the aggregation families and satisfied properties \cite{Tangwongsan-Sliding-Window-Aggregation-Algorithms}.

\begin{table}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/aggregation-operations-properties.png}
      \caption[Aggregation properties]{Aggregation properties. Checkmarks (\checkmark), crosses ($\times$), and question marks (?) indicate a property is true for all, false for all, or false for some of a given group of like operations, respectively}
      \label{tbl:aggregations-properties}
    \end{center}
\end{table}

@@

Based on the \textit{combine} and \textit{lower} functions, the following are formal definitions of these properties. An aggregation is said to have:

@@

The aggregation properties analyzed in this Section will be used to evaluate SWAG algorithms. Different algorithms might only apply to certain families of aggregations. This way, the multiple SWAG algorithms presented throughout this Thesis will be analyzed not only on their time and space complexity but also on the restrictions they impose regarding the aggregation' properties required. For example, under this framework, to implement Subtract-On-Evict (SOE), both the \textit{combine} $\oplus$ and its inverse $\ominus$ functions must be implemented so that SOE applies $\ominus$ to expire the evicted element from the aggregation state. Since SOE only restriction is for the aggregation to be invertible it works for the \textit{sum-like}, \textit{collect-like} and \textit{median-like} families in Table \ref{tbl:aggregations-properties}.

### 2020-02-24 chapters/sota.tex
Background edit. Two-Stacks.

In this Chapter, we will discuss the most relevant work for the aforementioned problem of detecting data pattern shifts in real-time computations over true sliding windows using resource-lightweight approaches on streaming systems. We present a wide analysis of several categories of algorithms that seemed like viable options to build the desired system.

\section{Detecting Change in Data Streams} \label{sec:change-detect}

For static data sets, it is reasonable to assume that the data was generated by a fixed process, for example, the data is a sample from a static distribution. But a data stream has necessarily a temporal dimension, and the underlying process that generates the data stream can change over time \cite{Aggarwal-Evolving-Data-Streams} \cite{Domingos-Mining-Time-Data-Streams}. It is this change in the underlying distribution that we want to detect and report as a data pattern shift.

\subsection*{A two window paradigm algorithm for change detection}
\label{subsec:2-window}

The authors of \cite{Kifer-Detecting-Change} propose an algorithm for change detection in data streams. The only assumption made is that data points are generated independently. The authors propose that detecting change in data streams can be reduced to testing if two windows have different underlying distributions. Thus the change detection algorithm works on a two-window paradigm.

Figure \ref{fig:change-detection-2-windows} contains the pseudocode presented by the authors. The proposed method uses two windows, a reference \textit{W\textsubscript{1}} and a sliding one \textit{W\textsubscript{2}}. The authors used tuple-based windows of size \textit{k} and the sliding window was a true sliding one, as described in Section \ref{sec:windows}. The reference window \textit{W\textsubscript{2}} works as a \textit{baseline} and contains the first \textit{k} points of the stream that occurred after the last detected change.

The algorithm begins by filling both windows with the first \textit{k} tuples. Then it slides window \textit{W\textsubscript{2}} (by one, since it is a true sliding window). It does this \textit{k} times. In between slides, the algorithm checks for change by applying the \textit{d} function, which measures the discrepancy between the two windows contents, \textit{W\textsubscript{1}} and \textit{W\textsubscript{2}}. Whenever this discrepancy is above a certain threshold $\alpha$, the algorithm reports a change in the data stream. Then it repeats the whole process, re-initializing both windows with the next \textit{k} items and proceeding as already described.

@@

This algorithm reduces change detection in data streams to testing whether two samples, the \textit{reference} and the \textit{sliding window} contents, are generated by different distributions. Consequently, the authors study the case of detecting differences in distribution between two samples. This is the purpose of the \textit{d} function. The authors evaluate several statistical methods as possible implementations of this \textit{d} function in order to encode an intuitive notion of change.

The following statistical tests for the \textit{d} function were used: Wilcoxon signed-rank, $\phi$\textsubscript{A} and $\Xi$\textsubscript{A}. The Wilcoxon test is well known but the $\phi$\textsubscript{A} and the $\Xi$\textsubscript{A} statistical tests were defined by the authors. Since computing these three statistical tests for our windows contents takes identical time and space, we consider that a thorough analysis of each one is out of scope for this Thesis, as it does not provide relevant information.

The authors conclude that no test is best in all of the experiments done. However, the $\phi$\textsubscript{A} and $\Xi$\textsubscript{A} statistics vastly outperformed the others in some tests and did not perform that much worse in other test cases.

Diving deeper in the algorithm presented in \ref{fig:change-detection-2-windows} we now analyze the computation of the \textit{d} function value. In the authors' work, efficient computation of the previously mentioned statistical tests --- \textit{i.e.} the value of function \textit{d} --- is further explored and an algorithm for such is presented. First, the \textit{KS structure} is presented. The formal definition given by the authors is shown in Figure \ref{fig:ks-structure}

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.4]{figures/ks-structure.png}
      \caption{Formal definition of the KS structure}
      \label{fig:ks-structure}
    \end{center}
\end{figure}

The KS structure is built from the reference \textit{W\textsubscript{1}} and sliding \textit{W\textsubscript{2}} windows. Consider the pair \textit{(\textit{W\textsubscript{1}}, \textit{W\textsubscript{2}})} where the size of both windows is \textit{k}. Creating the KS structure \textit{Z} of size \textit{2k} is done by joining the elements from \textit{W\textsubscript{1}} and \textit{W\textsubscript{2}}. As stated in \ref{fig:ks-structure}, each element of \textit{Z} will have a value and a weight. The values will be the actual elements joined. The weights of elements coming from \textit{W\textsubscript{1}} will be \textit{-1/k} while the weights of elements coming from \textit{W\textsubscript{2}} will be \textit{1/k}. The final KS structure \textit{Z} must be sorted in increasing order by value. The KS structure \textit{Z} can be maintained throughout running time by making use of a balanced tree, such as a B-tree, in \textit{O(log(2k))} time. By making use of the KS structure, all of the mentioned statistics can be computed in \textit{O(2k)} time.

\subsubsection*{Time complexity analysis}
%k to initialize and when change is detected
%k * d
%d 
 %nlog(2k) to initialize tree and when change is detected
 %log(2k) to maintain the tree
The algorithm begins by collecting the first \textit{k} tuples. Both windows will have size \textit{k}. This takes \textit{O(k)} time but is a task that will only occur on initialization and on change detection, upon which the window contents must reset, as described in \ref{fig:change-detection-2-windows}. For each of the next \textit{k} iterations the function \textit{d} will be computed and compared to the threshold $\alpha$. As previously mentioned, computing the value of \textit{d} can be done in \textit{O(log(2k))} time per iteration. However, the initial time cost of initializing the balanced tree is of \textit{O(k log(2k))}.

Initialization is not frequent. It occurs once at the beginning of execution and once per change detection upon which the algorithm resets the windows' contents. To provide an accurate time complexity, we define \textit{P} as the probability of a change occurring and being detected leading to a reset. Assuming \textit{P} is within $[0, 1]$, we have a time complexity of \textit{O(P(k + k log(2k)))} per reset --- \textit{i.e.} the probability of a reset occurring times \textit{O(k)} to initialize the windows' contents plus \textit{O(k log(2k))} to build the balanced tree.

In runtime, the algorithm will perform \textit{k} iterations where it computes the \textit{d} function value, taking \textit{O(log(2k))} per iteration, amounting to a time complexity of \textit{O(k log(2k)}.

Putting both initialization and runtime complexities we have a total time complexity of \textit{O(P(k + k log(2k)) + k log(2k))}.

\subsubsection*{Space complexity analysis}

The algorithm presented makes use of two \textit{k} sized windows and one balanced tree of size \textit{k} as well. This represents a space complexity of \textit{O(k+k+k)} or \textit{O(3k)}.

The algorithm for change detection presented in Subsection \ref{subsec:2-window} has time complexity of \textit{O(P(k + k log(2k)) + k log(2k))}. and space complexity of \textit{O(3k)}, with \textit{k} as a window size constant.

In this Thesis, we want to detect and alert data pattern shifts so detecting change in data streams is directly applicable to our use case. However, our focus is on a lightweight solution and this algorithm has a linear space complexity, so it does not fit our scope as a possible solution.

\section{Pattern Mining}

Taking on a new approach, in this Section, we explore pattern mining algorithms rather than algorithms that detect change in data streams. The idea behind using pattern mining algorithms being that if we can maintain a list of frequent patterns we would be able to detect when said patterns change --- \textit{i.e.} the list changes. Change in the frequent mined patterns would be associated with a data pattern shift and immediately reported.

\subsection{Mining Frequent Patterns: the FP-Stream structure}

@@

\subsubsection*{Applicability to our Hypothesis}
While a linear complexity fits most of the scenarios, it does not fit ours since we aim to build a low memory footprint solution, ideally with space complexity of \textit{O(1)}. Hence, this algorithm will be of no use to us.

@@

%\section{Sliding Window Aggregation Algorithms} \label{sec:sota-swag-algs}
%As discussed in Section \ref{sec:back-swag-algs} ...
%Besides RFS and SOE other methods were proposed like DABA...
%\subsection{DABA}
%\subsubsection*{Time complexity analysis}
%\subsubsection*{Space complexity analysis}
%\subsubsection*{Applicability to our Hypothesis}

\section{Probabilistic Data Structures} \label{sec:pds}
%TODO: UNCOMMENT
%In Section \ref{sec:sota-swag-algs} we presented some state-of-the-art algorithms for computing sliding window aggregations (SWAGs). However, if the aggregation we desire to compute can be approximated, there may exist more memory efficient methods. These aggregators are commonly denoted as sketches or probabilistic data structures and produce results orders-of magnitude faster then some exact approaches while keeping the memory consumption constant and providing mathematically proven error bounds. Each probabilistic data structure is used for efficient computation of one approximate aggregation, hence we say they are aggregation-specific.

@@

These aggregators are commonly denoted as sketches and produce results orders-of magnitude faster then some exact approaches while keeping the memory consumption constant and providing mathematically proven error bounds. Each probabilistic data structure is used for efficient computation of one approximate aggregation, hence we say they are aggregation-specific.

@@

\iffalse
\section{Sliding Window Aggregation with Probabilistic Data Structures}
In Section \ref{sec:pds} we presented a set of probabilistic data structures constant in time and space complexity. These structures aggregate an endless stream of data. However, when working under a sliding window context there is the need not only to insert new events but to evict old ones. Probabilistic data structures aggregate all incoming data but have no way of removing elements.

In this Section, we explore and analyze sliding window implementations of the previously discussed probabilistic data structures.

@@

\subsubsection{Time complexity analysis}

\subsubsection{Space complexity analysis}

\subsubsection{Applicability to our Hypothesis}

\fi

%\textbf{image for sliding hyperloglog}

>>>>

In this Chapter, we discuss the most relevant work for the aforementioned problem of detecting data pattern shifts in real-time computations over true sliding windows using resource-lightweight approaches on streaming systems. We present a wide analysis of several categories of algorithms that seemed like viable options to build the desired system.

\section{Data Mining} \label{sec:change-detect}

A data stream has a temporal dimension and the underlying process that generates the data stream can change over time \cite{Aggarwal-Evolving-Data-Streams} \cite{Domingos-Mining-Time-Data-Streams}. It is this change in the underlying distribution that we want to detect and report as a data pattern shift.

\subsection{A two window paradigm algorithm for change detection}
\label{subsec:2-window}

Kifer et al. propose an algorithm for change detection in data streams \cite{Kifer-Detecting-Change}. The authors propose that detecting change in data streams can be reduced to testing if two windows have different underlying distributions. Thus the change detection algorithm works on a two-window paradigm.

Figure \ref{fig:change-detection-2-windows} contains the pseudocode presented by the authors. The proposed method uses two windows, a reference \textit{W\textsubscript{1}} and a sliding one \textit{W\textsubscript{2}}. The authors use tuple-based windows of size \textit{k} and a true sliding window. The reference window \textit{W\textsubscript{2}} works as a \textit{baseline} and contains the first \textit{k} points of the stream that occurred after the last detected change.

The algorithm begins by filling both windows with the first \textit{k} tuples. Then it slides window \textit{W\textsubscript{2}} (by one tuple, since it is a true sliding window). It does this \textit{k} times and in between slides checks for change by applying the \textit{d} function, which measures the discrepancy between the two windows, \textit{W\textsubscript{1}} and \textit{W\textsubscript{2}}. Whenever this discrepancy is above a certain threshold $\alpha$, the algorithm reports a change in the data stream. Then it repeats the whole process, re-initializing both windows with the next \textit{k} items and proceeding as already described.

@@

This algorithm reduces change detection in data streams to testing whether two samples --- \textit{reference} and \textit{sliding} windows --- are generated by different distributions. The authors study methods for identifying differences in distribution between two samples and implement them as the \textit{d} function. Furthermore, they develop their own data structure and algorithm for efficient computation of \textit{d}.

%init - \textit{O(klog(k))}    
%runtime - \textit{O(klog(k)}

%k to initialize and when change is detected
%k * d
%d 
 %nlog(2k) to initialize tree and when change is detected
 %log(2k) to maintain the tree
\subsubsection*{Time complexity analysis}
Computing the value of \textit{d} can be done in \textit{O(log(k))} time per iteration. Hence, given \textit{k} iterations, the time complexity of this algorithm is \textit{O(k log(k))}.

\subsubsection*{Space complexity analysis}
The presented algorithm makes use of two \textit{k} sized windows and a custom data structure used for the efficient computation of \textit{d} of size \textit{k} as well. Since \textit{k} is constant, this method has constant space complexity.

\subsubsection*{Applicability to our Hypothesis}
In this Thesis, we want to detect and alert data pattern shifts so detecting change in data streams is directly applicable to our use case. However, our focus is on a lightweight solution and this algorithm has a linear space complexity, so it does not fit our scope as a possible solution.

%\subsection{Pattern Mining}

\subsection{Mining Frequent Patterns: the FP-Stream structure}
An alternative approach to detecting changes in data streams is to perform pattern mining. The idea behind using pattern mining algorithms being that if we can maintain a list of frequent patterns we would be able to detect when said patterns change --- \textit{i.e.} the list changes. Changes in the frequent mined patterns would be associated with a data pattern shift and immediately reported.

@@

\subsubsection*{Applicability to our Hypothesis}
While a linear complexity fits many use cases, it does not fit ours since we aim to build a low memory footprint solution, ideally with constant space complexity. Hence, this algorithm will be of no use to us.

@@

\section{Sliding Window Aggregation Algorithms} \label{sec:sota-swag-algs}

As discussed in Section \ref{sec:back-swag-algs}, a sliding window aggregation algorithm allows the computation of an exact or approximate aggregation over a sliding window. We have presented Recalculate-From-Scratch (RFS) and Subtract-On-Evict (SOE) as the two most basic generic stream processing algorithms. In this Section we analyze other more sophisticated SWAG algorithms that fit different use-cases, restraining the set of computable aggregations to certain families, as defined in Table \ref{tbl:aggregations-properties}.

\subsection{Two-Stacks}
The Two-Stacks sliding window aggregation (SWAG) algorithm works for scenarios where the sliding window has First-In First-Out (FIFO) semantics --- \textit{i.e.} elements are evicted in the same order they were inserted. Two-Stacks \cite{Tangwongsan-DABA} implements the FIFO queue using two stacks, namely the front stack F and the back stack B. Each element in each of the stacks is a pair \textit{<val, agg>} where \textit{val} is the item's value and \textit{agg} is the aggregation value of everything below it on the stack. The pseudocode for Two-Stacks \cite{Tangwongsan-DABA} can be seen in Figure \ref{fig:pseudo-2-stacks}.

Insertion of a new item is always made on stack B. The updated aggregation state is computed using the \textit{combine} function $\oplus$, taking as arguments the aggregation value \textit{agg} of the top element of stack B and the item's value \textit{val}. The pair pushed onto stack B is a pair with the item's value and the new aggregation value.

Evictions correspond to a simple pop from stack F. When F is empty, the elements from stack B are flipped onto stack F, re-applying $\oplus$ to make sure each element in F holds the aggregation value of everything below in the stack. Given \textit{n} as the size of stack B, the occasional flip operation takes \textit{O(n)} time.

Computing the sliding window aggregation is done by applying the \textit{combine} function $\oplus$ to the aggregation values of each of the stacks top element.

\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.45]{figures/2-stacks.png}
      \caption{Two-Stacks algorithm}
      \label{fig:pseudo-2-stacks}
    \end{center}
\end{figure}

\subsubsection*{Time complexity analysis}
Querying the aggregation state of the window and inserting a new element can be performed in constant time. However, while evicting an item is usually done in constant time, occasionally, a flip operation is required, consuming \textit{O(n)} time, with \textit{n} as the size of stack B. Because this flip is occasionally performed, the time complexity of Two-Stacks is said to be \textit{amortized} constant.

\subsubsection*{Space complexity analysis}
Two-Stacks uses two stacks as auxiliary data structures. The combined size of both stacks is at most the size of the sliding window \textit{n}. Hence, Two-Stacks has linear space complexity \textit{O(n)}.

\subsubsection*{Applicability to our Hypothesis}
Two-Stacks is a general sliding window aggregation algorithm to compute \textit{associative} aggregations. However, it is linear in space and occasionally performs the computationally intensive operation of flipping stack B onto stack F, making it unfit for our use case.

\subsection{DABA}
\subsubsection*{Time complexity analysis}
\subsubsection*{Space complexity analysis}
\subsubsection*{Applicability to our Hypothesis}

\section{Probabilistic Data Structures} \label{sec:pds}
In Section \ref{sec:sota-swag-algs} we presented some state-of-the-art algorithms for computing sliding window aggregations (SWAGs). However, if the aggregation we desire to compute can be approximated, there may exist more memory efficient methods. These aggregators are commonly denoted as sketches or probabilistic data structures and produce results orders-of magnitude faster then some exact approaches while keeping the memory consumption constant and providing mathematically proven error bounds. Each probabilistic data structure is used for efficient computation of one approximate aggregation, hence we say they are aggregation-specific.

@@

@@

\section{Sliding Window Aggregations with Probabilistic Data Structures}
In Section \ref{sec:pds} we presented a set of probabilistic data structures constant in time and space. These structures aggregate an endless stream of data. However, when working under a sliding window context there is the need not only to insert new events but to evict old ones. Probabilistic data structures aggregate all incoming data but have no way of expiring old elements from the aggregation state. In this Section, we explore sliding window implementations of the previously discussed probabilistic data structures.

\subsection{Sliding Bloom filter}
\subsubsection*{Time complexity analysis}
\subsubsection*{Space complexity analysis}
\subsubsection*{Applicability to our Hypothesis}

\subsection{Sliding Count-Min Sketch}
\subsubsection*{Time complexity analysis}
\subsubsection*{Space complexity analysis}
\subsubsection*{Applicability to our Hypothesis}

@@

\subsubsection*{Time complexity analysis}
\subsubsection*{Space complexity analysis}
\subsubsection*{Applicability to our Hypothesis}
